---
title: "160127-Multiple tests"
output: word_document

---

```{r,echo=FALSE}
# Umons _ Rapport d'analyse statistique automatisé pour dégrossissage des résultats de métaprotéomique et de qPCR. 
```

Strain : Sb

```{r,echo=FALSE}
require(car)

doss<-read.csv2("C:/Users/Utilisateur/Documents/R/MOOC_FUN_StatR/151105-Q-cm.csv",header=TRUE,sep=";",dec=".")


```
Summary

```{r,echo=FALSE}

summary(doss)

tab<-data.frame(table(doss$ASSAY),tapply(doss$QUANTITY,doss$ASSAY,mean),tapply(doss$QUANTITY,doss$ASSAY,sd))
tab<-tab[,2:4]
names(tab)<-c("Freq","Mean","Sd")
bou<-max(tab$Mean+2*tab$Sd)

tab1<-tab
tab1$Mean<-format(tab$Mean,scientific=TRUE,digits=3)
tab1$Sd<-format(tab$Sd,scientific=TRUE,digits=3)
tab1

par(mfcol=c(1,2))
boxplot(doss$QUANTITY~doss$ASSAY,xlab="Zinc concentration (mM)",ylab="Quantity (# ARN 16S)")
barplot(tab$Sd+tab$Mean,width=1,space=3,col="white")
#,ylim=c(0,bou))
barplot(tab$Mean,,width=1,space=3,col="black",names.arg=FALSE,add=TRUE)


```

Normality tests for Quantity and residus

```{r,echo=FALSE}

residu=doss$QUANTITY-ave(doss$QUANTITY)
par(mfcol=c(1,2))

qq.plot(doss$QUANTITY,distribution="norm",envelope=0.95,col="black",main="16S Quantity normality")
qq.plot(residu,distribution="norm",envelope=0.95,col="black",main="Residu normality")
```

Normality Tests: Shapiro-Wilk
If Shapiro-Wilk p-value > 0.05 => Normality, else, use Wilcoxon

```{r,echo=FALSE}

shapiro.test(doss$QUANTITY)
shapiro.test(residu)

```

Test d'homoscédasticité: Bartlett test
If p-value>0.05, then run ANOVA, else run Kruskal- Wallis

```{r,echo=FALSE}

bartlett.test(QUANTITY~ASSAY,data=doss)
```

Square deviation:
```{r,echo=FALSE}
tapply(doss$QUANTITY,doss$ASSAY,sd)

```

Statistical test: Anova

```{r,echo=FALSE}
lm(QUANTITY~ASSAY,data=doss)
drop1(lm(QUANTITY~ASSAY,data=doss),.~.,test="F")
aov(lm(QUANTITY~ASSAY,data=doss))
aov(QUANTITY~ASSAY,data=doss)
drop1(aov(QUANTITY~ASSAY,data=doss),.~.,test="F")
bou2<-aov(lm(QUANTITY~ASSAY,data=doss))
anova(aov(QUANTITY~ASSAY,data=doss))


wilcox.test(doss$QUANTITY~doss$ZINC)
library(stats)
TukeyHSD(lm(QUANTITY~ASSAY,data=doss))

```

Post-hoc ANOVA
Multiple T-test or Test de Tuckey ?

```{r,echo=FALSE}


pairwise.t.test(doss$QUANTITY,doss$ASSAY,p.adj="bonferroni",paired=FALSE)

library(stats)
anovov<-aov(QUANTITY~ASSAY,data=doss)
TukeyHSD(aov(QUANTITY~ASSAY,data=doss))


```

Statistical test: Kruskal- Wallis

```{r,echo=FALSE}
kruskal.test(doss$QUANTITY~doss$ASSAY)
```


Post-hoc Kruskal-Wallis
Dunn or Nemenyi ? and co test (by ranking)

```{r,echo=FALSE}
library(PMCMR)

posthoc.kruskal.nemenyi.test(doss$QUANTITY,doss$ASSAY,dist="Tukey")

#posthoc.kruskal.dunn.test(doss$QUANTITY,doss$ASSAY,p.adjust.method="bonferroni")
#same stuffs more or less... A bit less OK for some ^^


#dunn.test provides Dunn's (1964) z test approximation to a rank sum test employing both the same ranks used in the Kruskal-Wallis test, and the pooled variance estimate implied by the null hypothesis of the Kruskal-Wallis (akin to using the pooled variance to calculate t test statistics following an ANOVA).

#By contrast, the Kruskal-Nemenyi test as implemented in posthoc.kruskal.nemenyi.test is based on either the Studentized range distribution, or the ??2 distribution depending on user choice.

```


